---
title: "earthquake_API-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{earthquake_API-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Discription 
API for the USGS Earthquake Catalog ([link](https://earthquake.usgs.gov/fdsnws/event/1/)) that extracts recorded earthquakes that happend within a time period and that were of limited to a certain range of magnitude, both defined by the user.  

### The function accepts the following input from the user:
- **starttime**: Limit to events on or after the specified end start Enter character string "YYYY-MM-DD"  
- **endtime**: Limit to events on or before the specified end time. Enter character string "YYYY-MM-DD"  
- **min_magnitude**: Limit to events with a magnitude larger than the specified minimum, numeric  
- **max_magnitude**: Limit to events with a magnitude smaller than the specified maximum, numeric  

### Returns a data.frame containing:
- **time**: date (YYYY-MM-DD) and time event took place  
- **latitude**: recorded latitude of event  
- **longitude**: recorded longitude of event  
- **place**: Descriptive name of were event took place  

### Install package if needed 
```{r, eval=FALSE}
pak::pak("albba064/Lab5")
```

# Load package
```{r setup}
library(Lab5)
```

# Example 
## Make request
Get all earthquakes from USGS that happend between 25th of September 2026 and 1st of October 2026 that had a recorded magnitude 
between 5.5 and 10 
```{r}
output <- earthquake_API(starttime = "2026-09-25", 
                         endtime = "2026-10-01", 
                         min_magnitude = 5.5,
                         max_magnitude = 10)
```

## Output
Returns a data frame containing time, latitude, longitud and place for the observations
```{r}
knitr::kable(output)
```
